Unity in diversity: Children from 17 diverse communities process gaze in similar ways

### 2024-01-29 paper/paper.Rmd
update fig1

```{r fig1, include = T, fig.align = "center", fig.cap = "(B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting.", out.width="100%"}
knitr::include_graphics("../visuals/fig1.png")
```

>>>>

```{r fig1, include = T, fig.align = "center", fig.cap = "(A) Data collection sites. Points show approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting. (C) Graphical illustration of the cognitive model. Individuals infer the target of an agent's attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors. Individuals differ in their level of precision (indicated by alpha). Regardless of the level of precision, the further out the target lands, the less precise the model predicts individuals to be.", out.width="75%"}
knitr::include_graphics("../visuals/fig1.png")
```

### 2024-01-29 paper/paper.tex
update fig1

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Unity in diversity: Children from 17 communities process gaze in similar ways},
  pdfauthor={Manuel Bohn1,2, Julia Prein1,2, \ldots\ldots, \& Daniel B. M. Haun2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Unity in diversity: Children from 17 communities process gaze in similar ways}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{1,2}, \ldots{}\textsuperscript{\ldots{}}, \& Daniel B. M. Haun\textsuperscript{2}}
\date{}

@@

We conducted a large-scale cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures and their development. Previous developmental studies focused on the onset of gaze following in infancy (\emph{15}, \emph{16}). The XXXX participants (age ) in the study lived in 17 different communities spread over five continents. The countries from which data was contributed to the study represented 46\% of the world's population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (Fig. \ref{fig:fig1}).

@@

{\centering \includegraphics[width=1\linewidth]{../visuals/fig1}

@@

\caption{(B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting.}\label{fig:fig1}
\end{figure}

@@

We used an animated picture book tablet task in which participants were asked to locate a hidden object based on observing an agent's gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes. The key dependent variable was the precision with which children located the agent's focus of attention, that is, the deviation between where the agent looked (where the object was) and where the child thought the agent looked. We adapted visuals and audio instructions specifically to each of the 17 communities (Fig. 1). Previous work demonstrated excellent individual-level measurement properties for this task (\emph{17}). Thus, in addition to group-level trends, we were able to investigate individual-level variation.

As the first step, we investigated developmental gains. Across all 17 communities, we found a substantial increase in average levels of precision with age (fixed effect: \(\beta\) = -0.30, 95\% HDI (-0.39 - -0.21); range of community-level effects: \(\beta_{min}\) = -0.06, 95\% HDI (-0.17 - 0.06) to \(\beta_{max}\) = -0.58, 95\% HDI (-0.69 - -0.46). There were marked differences between communities: highly market-integrated communities around the globe showed higher levels of precision compared to less market-integrated communities. However, we think that these results are better explained by exposure to technology than market integration. We discuss them in more detail below. Differences between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Furthermore, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \ref{fig:fig2}A).

@@

Consistent developmental gains alone cannot inform us about the cognitive processes children use when locating the agent's focus of attention. Recent computational work modeled gaze following as social vector estimation (\emph{18}). When observing the eyes, onlookers estimate a vector from the center of the eye through the pupil. The focus of attention is the location wherever the estimated vectors from both eyes hit a surface (Fig. XX). The estimation process is not perfect, but each individual has a systematic level of uncertainty, which is conceptualized as the cause of individual differences. Importantly, this process model predicts a clear performance signature in our cross-cultural gaze following task: Trials in which the agent looks further away from the center (i.e.~to the left or right side of the screen) should result in lower levels of precision, compared to trials in which the agent looks closer to the middle. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual's level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area.

>>>>

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Unity in diversity: Children from 17 diverse communities process gaze in similar ways},
  pdfauthor={Manuel Bohn1,2, Julia Prein1,2, \ldots\ldots, \& Daniel B. M. Haun2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Unity in diversity: Children from 17 diverse communities process gaze in similar ways}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{1,2}, \ldots{}\textsuperscript{\ldots{}}, \& Daniel B. M. Haun\textsuperscript{2}}
\date{}

@@

We conducted a large-scale cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures and their development. Previous developmental studies focused on the onset of gaze following in infancy (\emph{15}, \emph{16}). The 1311 participants in the study lived in 17 different communities spread over five continents (Fig. \ref{fig:fig1}A, Tab. \ref{tab:tab1}). The countries from which data was contributed to the study represented 46 \% of the world's population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials).

@@

{\centering \includegraphics[width=0.75\linewidth]{../visuals/fig1}

@@

\caption{(A) Data collection sites. Points show approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting. (C) Graphical illustration of the cognitive model. Individuals infer the target of an agent's attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors. Individuals differ in their level of precision (indicated by alpha). Regardless of the level of precision, the further out the target lands, the less precise the model predicts individuals to be.}\label{fig:fig1}
\end{figure}

@@

We used an animated picture book tablet task in which participants were asked to locate a hidden object based on observing an agent's gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \ref{fig:fig1}B). The key dependent variable was the precision with which children located the agent's focus of attention, that is, the deviation between where the agent looked (where the object was) and where the child thought the agent looked. We adapted visuals and audio instructions specifically to each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task (\emph{17}). Thus, in addition to group-level trends, we were able to investigate individual-level variation.

As the first step, we investigated developmental gains. Across all 17 communities, we found a substantial increase in average levels of precision with age (fixed effect: \(\beta\) = -0.30, 95\% HDI (-0.39 - -0.21); range of community-level effects: \(\beta_{min}\) = -0.06, 95\% HDI (-0.17 - 0.06) to \(\beta_{max}\) = -0.58, 95\% HDI (-0.69 - -0.46). There were marked differences between communities: highly market-integrated communities around the globe showed higher levels of precision compared to less market-integrated communities. However, we believe that these results are better explained by exposure to technology than market integration. We discuss them in more detail below. Differences between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Furthermore, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \ref{fig:fig2}A).

@@

Consistent developmental gains alone cannot inform us about the cognitive processes children use when locating the agent's focus of attention. Recent computational work modeled gaze following as social vector estimation (\emph{18}). When observing the eyes, onlookers estimate a vector from the center of the eye through the pupil. The focus of attention is the location wherever the estimated vectors from both eyes hit a surface (Fig. \ref{fig:fig1}C). The estimation process is not perfect, but each individual has a systematic level of uncertainty, which is conceptualized as the cause of individual differences. Importantly, this process model predicts a clear performance signature in our cross-cultural gaze following task: Trials in which the agent looks further away from the center (i.e.~to the left or right side of the screen) should result in lower levels of precision, compared to trials in which the agent looks closer to the middle. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual's level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area.
